Using tours to visually investigate properties of new projection pursuit indexes with application to problems in physics

### 2019-08-06 scagnostics.Rmd
Outlining applications, filling in supplemental material section.

Three families of data simulations are used for examining the behaviour of the index functions. Each generates  structure in two variables, with the remaining variables containing various types of noise. This is a very simple construction, because there is no need for projection pursuit to find the structure, one could simply use the PPIs on pairs of variables. However, it serves the purpose to also evaluate the PPIs. The three data families are explained below. In each set, $n$ is used for the number of points, $p$ is the number of dimensions, and $d=2$ is the projection dimension. 
\textcolor{red}{The three structures were selected to cover both functional and non-functional dependence, different types of nuissance distributions and different structure size and squintability properties.}

@@

\textcolor{red}{For simplicity in the investigations of the index behavior, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution. }
The structured projection is in variables 5 and 6 ($x_5, x_6$). Two samples sizes are used: $n=(100, 1000)$. All variables are standardized to have mean 0 and standard deviation 1. Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables. Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000. The lower and upper show the 5th and 95th percentile of values. The holes index is sensitive only to the pipe distribution. All other indexes, except convex show distinctly higher values for the structured projections. The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs. The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix). Similarly, the TIC index is re-scaled depending on sample size.

@@

This section describes the application of these projection pursuit indices to find two-dimensional structure in two multidimensional gravitational waves problems.

@@

All of the indexes fall short against some aspect of the ideal properties of smoothness, squintability, flexibility, rotation invariance and speed. The paper describes how these properties can be assessed using tour methodology. Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
\textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes. It includes the functionalities needed to reproduce the assessements presented in this paper. While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.}

@@

- R code for paper
- spinebil
- galahr
- outline of appendix

>>>>

Three families of data simulations are used for examining the behaviour of the index functions. Each generates  structure in two variables, with the remaining variables containing various types of noise. This is a very simple construction, because there is no need for projection pursuit to find the structure, one could simply use the PPIs on pairs of variables. However, it serves the purpose to also evaluate the PPIs. The three data families are explained below. In each set, $n$ is used for the number of points, $p$ is the number of dimensions, and $d=2$ is the projection dimension. 
\textcolor{red}{The three structures were selected to cover both functional and non-functional dependence, different types of nuisance distributions and different structure size and squintability properties.}

@@

\textcolor{red}{For simplicity in the investigations of the index behaviour, we fix $p=6$, which corresponds two independent 2D planes containing nuisance distributions, and one 2D plane containing the structured distribution. }
The structured projection is in variables 5 and 6 ($x_5, x_6$). Two samples sizes are used: $n=(100, 1000)$. All variables are standardized to have mean 0 and standard deviation 1. Figure \ref{fig:data} shows samples from each of the families, of the nuisance and structured pairs of variables. Table \ref{tab:indexTable} compares the PPIs for structured projections against those for nuisance variables, based on 100 simulated data sets of each type, using sample size 1000. The lower and upper show the 5th and 95th percentile of values. The holes index is sensitive only to the pipe distribution. All other indexes, except convex show distinctly higher values for the structured projections. The convex index shows the inverse scale to other indices, thus (1-convex) will be used in the assessment of performance of PPIs. The scale for the holes index in its original implementation is smaller than the others ranging from about 0.7 through 1, so it is re-scaled in the performance assessment so that all indices can be plotted on a common scale of 0-1 (details in Appendix). Similarly, the TIC index is re-scaled depending on sample size.

@@

Applying projection pursuit to new datasets can be done using the guided tour available in the `tourr` package. The necessary steps are:

- Scaling and standardization of the data
- Index function selection and if necessary formatting according to the guided tour requirements, i.e. a function taking matrix input and returning the index value
- Calling the guided tour, for exploration this can be done via the `tourr::animate` function, for recording the results `tourr::save_history` should be used, examples are given in the documentation of the `tourr::guided_tour` function
- Things to explore: different indexes, different starting planes, different optimisation options, different ways of reducing dimensionality prior to the guided tour run

This section describes the application of these projection pursuit indices to find two-dimensional structure in two multidimensional gravitational waves problems.
\textcolor{red}{These illustrate different ways of exploring real data using projection pursuit, showcasing the dependence of the results on index function, starting plane and variable selection.}

@@

All of the indexes fall short against some aspect of the ideal properties of smoothness, squintability, flexibility, rotation invariance and speed. The paper describes how these properties can be assessed using tour methodology. Some potential fixes for the indexes are discussed but there is scope for further developing the new indexes.
\textcolor{red}{We recommend to use the spinebil [XXX REF] package when developing new indexes. It includes the functionalities needed to reproduce the assessments presented in this paper. While the current focus is on two-dimensional index functions, indexes in the tourr package apply to arbitrary projection dimension, and the methodology introduced here could be applied to the assessment of index functions where $d>2$.}

@@

This article was created with R Markdown [@rmarkdown], the code for the paper is available at [XXX].
The functions for testing new index functions as presented in this work are implemented in the R package spinebil [XXX].
The R package galahr [XXX] provides a graphical interface to the tourr package allowing for interactive exploration using the guided tour.
Additional information in the appendix is XXX
